Genetic draft and valley crossing

### 2018-05-29 ms/valley_crossing.tex
adding refs still...

\iffalse
(Note: There is an alternate regime in the appendix of \citet{good_2014} that would be better here. Will amend accordingly.)
We begin by considering the dynamics of $p_{\mathrm {fix}}$ before moving on to consider the distribution of bubble sizes.
To do so, we follow the method of \citet{good_2012} and \citet{good_2014}.
Let $p(x,s)$ be the fixation probability of the double mutant as a function of the absolute fitness $x$.
Furthermore, suppose that individuals experience ``background" mutations of small effect at one of a large number of loci.
These mutations occur at rate $U$ and are distributed as $\rho(\epsilon)$.
Then the fixation probability turns out to be governed by the integro-differential equation
\begin{equation}
v \partial_X p(X) = X p(X) + U \int \left[ p(X+\epsilon) - p(X)\right]\rho(\epsilon) d\epsilon - \frac{p(X)^2}{2},
\end{equation}
where $X = x-\sigma^2 t+s$ is the fitness relative to the (time dependent) mean fitness.

We zero in on the ``infinitesimal" model, where mutations are of small effect compared to $\sigma$.
In that case, the fixation probability is governed by the equation

\begin{equation}
v \partial_X p(X) = X p(X) + U \int \left[ p(X+\epsilon) - p(X)\right]\rho(\epsilon) d\epsilon - \frac{p(X)^2}{2},
\end{equation}

The perturbation provided by the mutation term is very important: without it, the equation is Bernoulli and can be solved readily, but the only solution is $p(X) = 0$ everywhere.
This should be intuitive, as lineages in a large population are doomed to extinction as the mean fitness inexorably advances, unless they an increase their fitness via recombination or mutation.
Assuming that $s$ is small compared to $\sigma$, a matched asymptotics analysis yields
\begin{equation}
p(X,s) =
\begin{cases}
0 &\mathrm{~if~} X+s < 0, \\
X_c e^{\frac{(X+s)^2-X_c^2}{2\sigma^2}} &\mathrm{~if~} 0 < X+s < X_c, \\
X+s &\mathrm{~if~} X+s > X_c.
\end{cases}
\end{equation}
(If $s$ is large compared to $\sigma$, then the standard Haldane result, $p(s) \sim s$, obtains.)
Effectively, this means the fixation probability is \emph{zero} if the double mutant lineage's fitness is less than the average for the population: if the fitness is less than some critical value $X_c$, then the lineage must generate a background mutant in order to fix, and if it is above $X_c$, the fixation probability is determined by the lineage's ability to survive drift and establish.
The value of $X_c$ is determined by the shape of $\rho(\epsilon)$.
Suppose $\rho$ is distributed exponentially: $\rho(\epsilon) = \frac{1}{\bar{\epsilon}} e^{-\epsilon/\bar{\epsilon}}$.
Then $X_c$ is determined by solving
\begin{equation}
2 = \frac{\sqrt{2\pi}\sigma U}{\bar{\epsilon}} \left[ \frac{1}{X_c - \frac{\sigma^2}{\bar{\epsilon}}} + \frac{\bar{\epsilon}}{\sigma^2} \left( 1+ \frac{\bar{\epsilon}}{X_c} \right) \right] e^{\frac{(X_c - \sigma^2/\bar{\epsilon})^2}{2\sigma^2}}. 
\end{equation}

We now consider the behavior of single mutant bubbles.
Since the tunneling probability for a lineage with background fitness $x$ is given by $\phi = \left< 1 - e^{-\mu p_{\mathrm fix} W} \right>$, we do not need to compute the full distribution of $W$ but rather only its Laplace transform: $\mathcal{L}\left[ p(w) \right] = \int_0^\infty e^{-zw} p(w) dw$, with $z = \mu p_{\mathrm {fix}}$.
We proceed to derive an equation for the time evolution of the Laplace transform.

Let $p(w,T|k,t,x)$ be the probability that a bubble has reached weight $w$ at time $T$, given that there were $n$ individuals present in the lineage at time $t$, and the bubble has background fitness $x$.
We have
\begin{align*}
-(\partial_t -k\partial_w) p(w,T|k,t,x) = &-k(2+x-\bar{x}-\delta)p(w,T|k,t,x) \\
& + k(1+x-\bar{x}-\delta)p(w,T|k+1,t,x) \\
& +kp(w,T|k-1,t,x).
\end{align*}
Consider the Laplace transform $\mathcal{L}\left[ p(w,T|k,t,x) \right] = \hat{p}(z,T|k,t,x) =  \int_0^\infty e^{-zw} p(w,T|k,t,x) dw$.
Using the convolution property of the transform, as well as the fact that each lineage is independent, allows us to write $\hat{p}(w,T|k,t,x) = \hat{p}^k(w,T|1,t,x)$.
We therefore have
\begin{equation}
-\partial_t\hat{p}(z,T|t,x) = -z\hat{p}(z,T|t,x) -(2+x-\bar{x}-\delta)\hat{p}(z,T|t,x) + (1+x-\bar{x}-\delta)\hat{p}(z,T|t,x)^2 + 1.
\end{equation}
Defining $\phi = 1-\hat{p}$ (consistent with our previous definition of $\phi$ and normalizing our variables via $\tau = \sigma(T-t)$, $\chi = x/\sigma-\sigma T$, $\tilde{z} = z/\sigma$, and $\tilde{\delta} = \delta/\sigma$ yields
\begin{equation}
\partial_\tau \tilde{\phi} = \tilde{z} + (\tilde{z} + \chi + \tau + \tilde{\delta})\tilde{\phi} - (1 + \chi + \tau + \tilde{\delta})\tilde{\phi}^2,
\end{equation}
where $\tilde{\phi}$ = $\phi/\sigma$.
We assume that the prefactor of $\tilde{\phi}^2$ is approximately $1$, which corresponds to the growth rate being small compared to $1$.
In addition, we will eventually want to take the limit as $\tau \to \infty$, which corresponds to averaging over all possible bubble sizes, no matter how long lived they are.
However, this will yield the absurd result that $\tilde{\phi}$ grows without bound.
To remedy this, we assume that the lifespan of bubbles is short compared to the time scale over which selection acts, meaning that we drop the $\tau$ term.
Defining $\theta = \tilde{z} + \chi + \tilde{\delta}$ yields
\begin{equation}
\partial_\tau \tilde{\phi} = \tilde{z} + \theta \tilde{\phi} - \tilde{\phi}^2.
\end{equation}
Using matched asymptotics allows us to write
\begin{equation}
\tilde{\phi} = 
\begin{cases}
\tilde{z} &\mathrm{~if~} \theta < 0, \\
\theta &\mathrm{~if~} \theta > 0.
\end{cases}
\end{equation}
Intuitively, this corresponds to stating that bubbles with a positive growth rate are likely to persist for some time, but bubbles with a negative growth rate are doomed to extinction: the only way they can succeed is to generate a successful double mutant in $O(1)$ generations.
Knowing the form of $\phi$ and $z$ allows us to compute the overall tunneling probability and hence the expected time to a tunneling event, once we integrate over the fitness distribution $f(x)$.
\fi

@@

\iffalse
\begin{figure}
\includegraphics[width=0.4\textwidth]{Figures/bands.pdf}
\caption{Valley crossing times for fixed $N\sigma$, with $N$ allowed to vary up or down by a factor of $2$ (and $\sigma$ adjusted accordingly). Error bars are quartile ranges. Here, $\mu = 3 \times 10^{-6}$, $\delta = 10^{-3}$, and $s = 10^{-2}$.}
\label{fig:bands}
\end{figure}

In addition, we consider the effects of the compound parameter $N\sigma$, which has previously been shown to determine the coalescent behavior of populations \citep{neher_hallatschek_2013, neher_kessinger_2013}.
Figure \ref{fig:bands} suggests that, at high values of $\sigma$, the value of $N\sigma$ is more important than $N$ itself in determining the valley crossing rate.
The dependence on $N$ appears to be reduced, consistent with general features of very large, rapidly adapting populations.
In a slowly adapting population, $N$ determines the transition from sequential fixation to tunneling: if bubbles reach a maximum size that is of order $N$, they start to noticeably decrease the mean fitness, making fixation easier.
When $N\sigma \gg 1$, bubbles almost never reach a large enough size for this condition to obtain.
\fi

@@

\iffalse
First, we consider the case where $\rho = 0$ and focus on what \citet{weissman_2009} refer to as the process of ``stochastic tunneling", where $N\mu \ll 1$.
As we shall see, the results in the $\rho = 0$ and low $\rho$ limits can easily be extended to sexually recombining populations: moreover, understanding the rate of tunneling will allow us to easily examine the rate at which sequential fixation and deterministic fixation occur.
In stochastic tunneling, a deleterious intermediate arises (at rate $N\mu$), and its lineage persists over a short time $T$, forming a transient ``bubble".
If the number of individuals in the lineage is given by $n(t)$, then the total number of individuals in the ``bubble" is $w = \int_{t_0}^T n(t) dt$.
We refer to $w$ as the ``weight" of the bubble.
The overall rate of valley crossing becomes $1/N\mu p$, with $p$ the probability that a particular bubble will give rise to a successful double mutant.
Thus, given that a deleterious intermediate appears, it has $w$ ``chances" to generate a successful complex adaptation, which it does at rate $\sim \mu s$: double mutants arise at rate $\mu$ and have a probability $\sim s$ of fixing.
The probability of achieving a weight of at least $w$ scales as $1/\sqrt{w}$ (or, equivalently, $P(w) \sim w^{-3/2}$), meaning that with probability $\sqrt{\mu s}$, a bubble gives rise to a successful single mutant.
This probability drops off rapidly at $1/\delta^2$, so for higher values of $\delta$ (so called ``deleterious tunneling"), the rate at which new complex adaptations are generated starts to fall: the size of the mutant subpopulation is bounded from above by roughly $1/\delta$.

In the high $N\sigma$ regime, the argument is much the same, but the underlying dynamics are different.
In stark contrast to the drift case, the fate of a complex adaptation depends strongly on the fitness $x$ of the genetic background on which it appears.
Individuals with fitness lower than some critical fitness $x_c$ are almost guaranteed to go extinct.
This means the fixation probability of the complex adaptation no longer scales with $s$, and $P(w) \sim w^{-2}$ instead of $w^{-3/2}$, meaning that the probability of reaching weight $w$ is $\sim 1/w$.
However, individuals with higher fitnesses can be destined for fixation, even if they are carrying a significant load of deleterious mutations.
This means the dependence on $\delta$ is significantly lessened: even deep fitness valleys can be crossed by a rapid enough wave.
(Still needed: expression for fixation probability of beneficial mutant--this should be easy enough to obtain.
Can start with equation 3 of \citet{neher_shraiman_2011} and drop the recombination term, which is what I'm doing now.
This is tantamount to a ``mean field" approximation but might still work out okay.
Other possible approach is to use something like the expression of  \citet{desai_fisher_2007} for the fixation probability and rescale it to a traveling wave model.
Once we have the fixation probability, studying sequential and deterministic fixation will be easier than studying tunneling, since sequential fixation is essentially two different fixation events.)

One difficulty in computing the rate of tunneling in the high $N\sigma$ case is that the expected bubble size and fixation probability will both turn out to depend on the background fitness of an individual.
That is to say, the total rate of valley crossing will be given by $\int P(x) p_{\mathrm{fix}}(x,s) \mathbb{E}\left[ w(x,\delta) \right] dx$, with $P(x)$ the (presumably Gaussian) background fitness distribution.
One cannot simply compute $p_{\mathrm{fix}}$ and the weight distribution by themselves but rather must convolute both of them over the background fitness.
This is intuitive: individuals near the nose of the distribution will tend to give rise to longer-lived lineages, and because they are near the nose, the complex adaptation will therefore likewise be more likely to fix.
We proceed to derive both $p_{\mathrm{fix}}(x,s)$ and $P(w)$ in the zero recombination limit.

The establishment probability $p(x,s)$ obeys the PDE
\begin{equation}
v \partial_x p(x,s) = (x+s)p(x,s) - (1+x+s)p(x,s)^2.
\end{equation}
To see this, consider the generic expression for the fixation probability
\begin{equation}
p(X,t-dt) = p(X,t)-dt(D+B(X,t))p(X,t) + dt(B(X,t)(2p(X,t)-p(X,t)^2)
\end{equation}
\citep{barton_1995}. Taking the limit as $dt \to 0$ and inserting $D = 1, B = 1 + X -\bar{X}(t)+s$ yields a PDE for $p(X,t)$.
Defining $x = X - vt$, we then have $\partial_t p(X-vt) = -v\partial_x p(x)$.

This is a Bernoulli equation and is readily solved by substituting $u(x,s) = 1/p(x,s)$. We then have that $\partial_x u(x,s) = -\frac{\partial_x p(x,s)}{p(x,s)^2}$, so that
\begin{equation}
\partial_x u(x,s) = -\frac{(x+s)}{v}u(x,s) - \frac{1+x+s}{v}.
\end{equation}
The use of the integrating factor $e^{\int \frac{x+s}{v} dx} = Ce^{\frac{(x+s)^2}{2v}}$ yields
\begin{equation}
Ce^{\frac{(x+s)^2}{2v}} (\partial_x u + \frac{x+s}{v}u) = -Ce^{\frac{(x+s)^2}{2v}}\frac{1+x+s}{v}.
\end{equation}
The left side is simply $\partial_x(Ce^{\frac{(x+s)^2}{2v}}u)$, so (after dividing $C$ out) we can integrate both sides, then multiply by $e^{\frac{-(x+s)^2}{2v}}$ to obtain
\begin{equation}
u(x,s) =  -e^{\frac{-(x+s)^2}{2v}} \frac{\sqrt{\pi /2} \textrm{erfi}(\frac{x+s}{\sqrt{2v}}}{\sqrt{v}}) - 1 + C(e^{\frac{-(x+s)^2}{2v}}),
\end{equation}
with $C$ a constant. Of course, $p = 1/u$.

Unfortunately it is not clear how to get rid of $C$, as there is no obvious ``initial condition".
The most we are likely to know about $p$ is that $p$ approaches $0$ in the limit of low $x$, but behaves $\sim s$ in the limit of high $x$ before saturating at $1$ (however, there are unlikely to be any individuals in the population at such high $x$).
One possibility is to impose an \emph{ersatz} ``solvability condition": surely the fixation probability of a neutral mutation averaged over the entire population must equal $1/N$. We would then have
\begin{equation}
\int_{-\infty}^{\infty} P(x)p(x,0)dx = \frac{1}{\sqrt{2\pi}\sigma}\int_{-\infty}^{\infty} \frac{e^{-x^2/2\sigma^2}}{-e^{-x^2/2\sigma^2} \frac{\sqrt{\pi /2} \textrm{erfi}(\frac{x}{\sqrt{2\sigma^2}})}{\sigma} - 1 + Ce^{-x^2/2\sigma^2}}dx = \frac{1}{N}.
\end{equation}
Multiplying the top and bottom of the integrand by $e^{x^2/2\sigma^2}$ would clean this up, yielding
\begin{equation}
\frac{1}{\sqrt{2\pi}\sigma}\int_{-\infty}^{\infty} \frac{1}{-\frac{\sqrt{\pi /2} \textrm{erfi}(\frac{x}{\sqrt{2\sigma^2}})}{\sigma} - e^{x^2/2\sigma^2} + C}dx = \frac{1}{N}.
\end{equation}
Unfortunately this integral has an imaginary error function in the denominator, so there is not likely to be a closed form expression: we could explore it numerically and see where it peaks, however.

Another idea would be to consider an alternative way of solving the differential equation for $p$.
Return to
\begin{equation}
v \partial_x p(x,s) = (x+s)p(x,s) - (1+x+s)p(x,s)^2.
\end{equation}
Suppose we make the \emph{ansatz} that $x+s \ll 1$, and define $\chi = x/\sigma$, $\tilde{s} = s/\sigma$, and $\tilde{p}(\chi,\tilde{s}) = p(x,s)/\sigma$. Then
\begin{equation}
\partial_\chi \tilde{p}(\chi,\tilde{s}) = (\chi + \tilde{s})\tilde{p}(\chi,\tilde{s}) - \tilde{p}(\chi,\tilde{s})^2.
\end{equation}
Now define $q(\chi,\tilde{s})$ such that $\tilde{w}(\chi,\tilde{s}) = \partial_x \log q(\chi,\tilde{s})$.
Then
\begin{equation}
\partial_\chi^2 q(\chi,\tilde{s}) - (\chi+\tilde{s})\partial_\chi q(\chi, \tilde{s}) = 0.
\end{equation}
Defining $\vartheta = \chi + \tilde{s}$ and $q(\vartheta) = e^{\vartheta^2/4}h(\vartheta)$ yields
\begin{equation}
\partial_\vartheta^2 h(\vartheta) - (\frac{1}{2} + \frac{\theta^2}{4})h(\vartheta) = 0,
\end{equation}
which is the parabolic cylinder equation.
We then need to find the form of $q$ that has the correct asymptotics.
This does have a solution:
\begin{equation}
q(\chi) = \int_0^\infty d\lambda e^{\vartheta \lambda - \lambda^2/2}/\lambda.
\end{equation}

>>>>

% \iffalse
% (Note: There is an alternate regime in the appendix of \citet{good_2014} that would be better here. Will amend accordingly.)
% We begin by considering the dynamics of $p_{\mathrm {fix}}$ before moving on to consider the distribution of bubble sizes.
% To do so, we follow the method of \citet{good_2012} and \citet{good_2014}.
% Let $p(x,s)$ be the fixation probability of the double mutant as a function of the absolute fitness $x$.
% Furthermore, suppose that individuals experience ``background" mutations of small effect at one of a large number of loci.
% These mutations occur at rate $U$ and are distributed as $\rho(\epsilon)$.
% Then the fixation probability turns out to be governed by the integro-differential equation
% \begin{equation}
% v \partial_X p(X) = X p(X) + U \int \left[ p(X+\epsilon) - p(X)\right]\rho(\epsilon) d\epsilon - \frac{p(X)^2}{2},
% \end{equation}
% where $X = x-\sigma^2 t+s$ is the fitness relative to the (time dependent) mean fitness.

% We zero in on the ``infinitesimal" model, where mutations are of small effect compared to $\sigma$.
% In that case, the fixation probability is governed by the equation

% \begin{equation}
% v \partial_X p(X) = X p(X) + U \int \left[ p(X+\epsilon) - p(X)\right]\rho(\epsilon) d\epsilon - \frac{p(X)^2}{2},
% \end{equation}

% The perturbation provided by the mutation term is very important: without it, the equation is Bernoulli and can be solved readily, but the only solution is $p(X) = 0$ everywhere.
% This should be intuitive, as lineages in a large population are doomed to extinction as the mean fitness inexorably advances, unless they an increase their fitness via recombination or mutation.
% Assuming that $s$ is small compared to $\sigma$, a matched asymptotics analysis yields
% \begin{equation}
% p(X,s) =
% \begin{cases}
% 0 &\mathrm{~if~} X+s < 0, \\
% X_c e^{\frac{(X+s)^2-X_c^2}{2\sigma^2}} &\mathrm{~if~} 0 < X+s < X_c, \\
% X+s &\mathrm{~if~} X+s > X_c.
% \end{cases}
% \end{equation}
% (If $s$ is large compared to $\sigma$, then the standard Haldane result, $p(s) \sim s$, obtains.)
% Effectively, this means the fixation probability is \emph{zero} if the double mutant lineage's fitness is less than the average for the population: if the fitness is less than some critical value $X_c$, then the lineage must generate a background mutant in order to fix, and if it is above $X_c$, the fixation probability is determined by the lineage's ability to survive drift and establish.
% The value of $X_c$ is determined by the shape of $\rho(\epsilon)$.
% Suppose $\rho$ is distributed exponentially: $\rho(\epsilon) = \frac{1}{\bar{\epsilon}} e^{-\epsilon/\bar{\epsilon}}$.
% Then $X_c$ is determined by solving
% \begin{equation}
% 2 = \frac{\sqrt{2\pi}\sigma U}{\bar{\epsilon}} \left[ \frac{1}{X_c - \frac{\sigma^2}{\bar{\epsilon}}} + \frac{\bar{\epsilon}}{\sigma^2} \left( 1+ \frac{\bar{\epsilon}}{X_c} \right) \right] e^{\frac{(X_c - \sigma^2/\bar{\epsilon})^2}{2\sigma^2}}. 
% \end{equation}

% We now consider the behavior of single mutant bubbles.
% Since the tunneling probability for a lineage with background fitness $x$ is given by $\phi = \left< 1 - e^{-\mu p_{\mathrm fix} W} \right>$, we do not need to compute the full distribution of $W$ but rather only its Laplace transform: $\mathcal{L}\left[ p(w) \right] = \int_0^\infty e^{-zw} p(w) dw$, with $z = \mu p_{\mathrm {fix}}$.
% We proceed to derive an equation for the time evolution of the Laplace transform.

% Let $p(w,T|k,t,x)$ be the probability that a bubble has reached weight $w$ at time $T$, given that there were $n$ individuals present in the lineage at time $t$, and the bubble has background fitness $x$.
% We have
% \begin{align*}
% -(\partial_t -k\partial_w) p(w,T|k,t,x) = &-k(2+x-\bar{x}-\delta)p(w,T|k,t,x) \\
% & + k(1+x-\bar{x}-\delta)p(w,T|k+1,t,x) \\
% & +kp(w,T|k-1,t,x).
% \end{align*}
% Consider the Laplace transform $\mathcal{L}\left[ p(w,T|k,t,x) \right] = \hat{p}(z,T|k,t,x) =  \int_0^\infty e^{-zw} p(w,T|k,t,x) dw$.
% Using the convolution property of the transform, as well as the fact that each lineage is independent, allows us to write $\hat{p}(w,T|k,t,x) = \hat{p}^k(w,T|1,t,x)$.
% We therefore have
% \begin{equation}
% -\partial_t\hat{p}(z,T|t,x) = -z\hat{p}(z,T|t,x) -(2+x-\bar{x}-\delta)\hat{p}(z,T|t,x) + (1+x-\bar{x}-\delta)\hat{p}(z,T|t,x)^2 + 1.
% \end{equation}
% Defining $\phi = 1-\hat{p}$ (consistent with our previous definition of $\phi$ and normalizing our variables via $\tau = \sigma(T-t)$, $\chi = x/\sigma-\sigma T$, $\tilde{z} = z/\sigma$, and $\tilde{\delta} = \delta/\sigma$ yields
% \begin{equation}
% \partial_\tau \tilde{\phi} = \tilde{z} + (\tilde{z} + \chi + \tau + \tilde{\delta})\tilde{\phi} - (1 + \chi + \tau + \tilde{\delta})\tilde{\phi}^2,
% \end{equation}
% where $\tilde{\phi}$ = $\phi/\sigma$.
% We assume that the prefactor of $\tilde{\phi}^2$ is approximately $1$, which corresponds to the growth rate being small compared to $1$.
% In addition, we will eventually want to take the limit as $\tau \to \infty$, which corresponds to averaging over all possible bubble sizes, no matter how long lived they are.
% However, this will yield the absurd result that $\tilde{\phi}$ grows without bound.
% To remedy this, we assume that the lifespan of bubbles is short compared to the time scale over which selection acts, meaning that we drop the $\tau$ term.
% Defining $\theta = \tilde{z} + \chi + \tilde{\delta}$ yields
% \begin{equation}
% \partial_\tau \tilde{\phi} = \tilde{z} + \theta \tilde{\phi} - \tilde{\phi}^2.
% \end{equation}
% Using matched asymptotics allows us to write
% \begin{equation}
% \tilde{\phi} = 
% \begin{cases}
% \tilde{z} &\mathrm{~if~} \theta < 0, \\
% \theta &\mathrm{~if~} \theta > 0.
% \end{cases}
% \end{equation}
% Intuitively, this corresponds to stating that bubbles with a positive growth rate are likely to persist for some time, but bubbles with a negative growth rate are doomed to extinction: the only way they can succeed is to generate a successful double mutant in $O(1)$ generations.
% Knowing the form of $\phi$ and $z$ allows us to compute the overall tunneling probability and hence the expected time to a tunneling event, once we integrate over the fitness distribution $f(x)$.
% \fi

@@

% \iffalse
% \begin{figure}
% \includegraphics[width=0.4\textwidth]{Figures/bands.pdf}
% \caption{Valley crossing times for fixed $N\sigma$, with $N$ allowed to vary up or down by a factor of $2$ (and $\sigma$ adjusted accordingly). Error bars are quartile ranges. Here, $\mu = 3 \times 10^{-6}$, $\delta = 10^{-3}$, and $s = 10^{-2}$.}
% \label{fig:bands}
% \end{figure}

% In addition, we consider the effects of the compound parameter $N\sigma$, which has previously been shown to determine the coalescent behavior of populations \citep{neher_hallatschek_2013, neher_kessinger_2013}.
% Figure \ref{fig:bands} suggests that, at high values of $\sigma$, the value of $N\sigma$ is more important than $N$ itself in determining the valley crossing rate.
% The dependence on $N$ appears to be reduced, consistent with general features of very large, rapidly adapting populations.
% In a slowly adapting population, $N$ determines the transition from sequential fixation to tunneling: if bubbles reach a maximum size that is of order $N$, they start to noticeably decrease the mean fitness, making fixation easier.
% When $N\sigma \gg 1$, bubbles almost never reach a large enough size for this condition to obtain.
% \fi

@@

% \iffalse
% First, we consider the case where $\rho = 0$ and focus on what \citet{weissman_2009} refer to as the process of ``stochastic tunneling", where $N\mu \ll 1$.
% As we shall see, the results in the $\rho = 0$ and low $\rho$ limits can easily be extended to sexually recombining populations: moreover, understanding the rate of tunneling will allow us to easily examine the rate at which sequential fixation and deterministic fixation occur.
% In stochastic tunneling, a deleterious intermediate arises (at rate $N\mu$), and its lineage persists over a short time $T$, forming a transient ``bubble".
% If the number of individuals in the lineage is given by $n(t)$, then the total number of individuals in the ``bubble" is $w = \int_{t_0}^T n(t) dt$.
% We refer to $w$ as the ``weight" of the bubble.
% The overall rate of valley crossing becomes $1/N\mu p$, with $p$ the probability that a particular bubble will give rise to a successful double mutant.
% Thus, given that a deleterious intermediate appears, it has $w$ ``chances" to generate a successful complex adaptation, which it does at rate $\sim \mu s$: double mutants arise at rate $\mu$ and have a probability $\sim s$ of fixing.
% The probability of achieving a weight of at least $w$ scales as $1/\sqrt{w}$ (or, equivalently, $P(w) \sim w^{-3/2}$), meaning that with probability $\sqrt{\mu s}$, a bubble gives rise to a successful single mutant.
% This probability drops off rapidly at $1/\delta^2$, so for higher values of $\delta$ (so called ``deleterious tunneling"), the rate at which new complex adaptations are generated starts to fall: the size of the mutant subpopulation is bounded from above by roughly $1/\delta$.

% In the high $N\sigma$ regime, the argument is much the same, but the underlying dynamics are different.
% In stark contrast to the drift case, the fate of a complex adaptation depends strongly on the fitness $x$ of the genetic background on which it appears.
% Individuals with fitness lower than some critical fitness $x_c$ are almost guaranteed to go extinct.
% This means the fixation probability of the complex adaptation no longer scales with $s$, and $P(w) \sim w^{-2}$ instead of $w^{-3/2}$, meaning that the probability of reaching weight $w$ is $\sim 1/w$.
% However, individuals with higher fitnesses can be destined for fixation, even if they are carrying a significant load of deleterious mutations.
% This means the dependence on $\delta$ is significantly lessened: even deep fitness valleys can be crossed by a rapid enough wave.
% (Still needed: expression for fixation probability of beneficial mutant--this should be easy enough to obtain.
% Can start with equation 3 of \citet{neher_shraiman_2011} and drop the recombination term, which is what I'm doing now.
% This is tantamount to a ``mean field" approximation but might still work out okay.
% Other possible approach is to use something like the expression of  \citet{desai_fisher_2007} for the fixation probability and rescale it to a traveling wave model.
% Once we have the fixation probability, studying sequential and deterministic fixation will be easier than studying tunneling, since sequential fixation is essentially two different fixation events.)

% One difficulty in computing the rate of tunneling in the high $N\sigma$ case is that the expected bubble size and fixation probability will both turn out to depend on the background fitness of an individual.
% That is to say, the total rate of valley crossing will be given by $\int P(x) p_{\mathrm{fix}}(x,s) \mathbb{E}\left[ w(x,\delta) \right] dx$, with $P(x)$ the (presumably Gaussian) background fitness distribution.
% One cannot simply compute $p_{\mathrm{fix}}$ and the weight distribution by themselves but rather must convolute both of them over the background fitness.
% This is intuitive: individuals near the nose of the distribution will tend to give rise to longer-lived lineages, and because they are near the nose, the complex adaptation will therefore likewise be more likely to fix.
% We proceed to derive both $p_{\mathrm{fix}}(x,s)$ and $P(w)$ in the zero recombination limit.

% The establishment probability $p(x,s)$ obeys the PDE
% \begin{equation}
% v \partial_x p(x,s) = (x+s)p(x,s) - (1+x+s)p(x,s)^2.
% \end{equation}
% To see this, consider the generic expression for the fixation probability
% \begin{equation}
% p(X,t-dt) = p(X,t)-dt(D+B(X,t))p(X,t) + dt(B(X,t)(2p(X,t)-p(X,t)^2)
% \end{equation}
% \citep{barton_1995}. Taking the limit as $dt \to 0$ and inserting $D = 1, B = 1 + X -\bar{X}(t)+s$ yields a PDE for $p(X,t)$.
% Defining $x = X - vt$, we then have $\partial_t p(X-vt) = -v\partial_x p(x)$.

% This is a Bernoulli equation and is readily solved by substituting $u(x,s) = 1/p(x,s)$. We then have that $\partial_x u(x,s) = -\frac{\partial_x p(x,s)}{p(x,s)^2}$, so that
% \begin{equation}
% \partial_x u(x,s) = -\frac{(x+s)}{v}u(x,s) - \frac{1+x+s}{v}.
% \end{equation}
% The use of the integrating factor $e^{\int \frac{x+s}{v} dx} = Ce^{\frac{(x+s)^2}{2v}}$ yields
% \begin{equation}
% Ce^{\frac{(x+s)^2}{2v}} (\partial_x u + \frac{x+s}{v}u) = -Ce^{\frac{(x+s)^2}{2v}}\frac{1+x+s}{v}.
% \end{equation}
% The left side is simply $\partial_x(Ce^{\frac{(x+s)^2}{2v}}u)$, so (after dividing $C$ out) we can integrate both sides, then multiply by $e^{\frac{-(x+s)^2}{2v}}$ to obtain
% \begin{equation}
% u(x,s) =  -e^{\frac{-(x+s)^2}{2v}} \frac{\sqrt{\pi /2} \textrm{erfi}(\frac{x+s}{\sqrt{2v}}}{\sqrt{v}}) - 1 + C(e^{\frac{-(x+s)^2}{2v}}),
% \end{equation}
% with $C$ a constant. Of course, $p = 1/u$.

% Unfortunately it is not clear how to get rid of $C$, as there is no obvious ``initial condition".
% The most we are likely to know about $p$ is that $p$ approaches $0$ in the limit of low $x$, but behaves $\sim s$ in the limit of high $x$ before saturating at $1$ (however, there are unlikely to be any individuals in the population at such high $x$).
% One possibility is to impose an \emph{ersatz} ``solvability condition": surely the fixation probability of a neutral mutation averaged over the entire population must equal $1/N$. We would then have
% \begin{equation}
% \int_{-\infty}^{\infty} P(x)p(x,0)dx = \frac{1}{\sqrt{2\pi}\sigma}\int_{-\infty}^{\infty} \frac{e^{-x^2/2\sigma^2}}{-e^{-x^2/2\sigma^2} \frac{\sqrt{\pi /2} \textrm{erfi}(\frac{x}{\sqrt{2\sigma^2}})}{\sigma} - 1 + Ce^{-x^2/2\sigma^2}}dx = \frac{1}{N}.
% \end{equation}
% Multiplying the top and bottom of the integrand by $e^{x^2/2\sigma^2}$ would clean this up, yielding
% \begin{equation}
% \frac{1}{\sqrt{2\pi}\sigma}\int_{-\infty}^{\infty} \frac{1}{-\frac{\sqrt{\pi /2} \textrm{erfi}(\frac{x}{\sqrt{2\sigma^2}})}{\sigma} - e^{x^2/2\sigma^2} + C}dx = \frac{1}{N}.
% \end{equation}
% Unfortunately this integral has an imaginary error function in the denominator, so there is not likely to be a closed form expression: we could explore it numerically and see where it peaks, however.

% Another idea would be to consider an alternative way of solving the differential equation for $p$.
% Return to
% \begin{equation}
% v \partial_x p(x,s) = (x+s)p(x,s) - (1+x+s)p(x,s)^2.
% \end{equation}
% Suppose we make the \emph{ansatz} that $x+s \ll 1$, and define $\chi = x/\sigma$, $\tilde{s} = s/\sigma$, and $\tilde{p}(\chi,\tilde{s}) = p(x,s)/\sigma$. Then
% \begin{equation}
% \partial_\chi \tilde{p}(\chi,\tilde{s}) = (\chi + \tilde{s})\tilde{p}(\chi,\tilde{s}) - \tilde{p}(\chi,\tilde{s})^2.
% \end{equation}
% Now define $q(\chi,\tilde{s})$ such that $\tilde{w}(\chi,\tilde{s}) = \partial_x \log q(\chi,\tilde{s})$.
% Then
% \begin{equation}
% \partial_\chi^2 q(\chi,\tilde{s}) - (\chi+\tilde{s})\partial_\chi q(\chi, \tilde{s}) = 0.
% \end{equation}
% Defining $\vartheta = \chi + \tilde{s}$ and $q(\vartheta) = e^{\vartheta^2/4}h(\vartheta)$ yields
% \begin{equation}
% \partial_\vartheta^2 h(\vartheta) - (\frac{1}{2} + \frac{\theta^2}{4})h(\vartheta) = 0,
% \end{equation}
% which is the parabolic cylinder equation.
% We then need to find the form of $q$ that has the correct asymptotics.
% This does have a solution:
% \begin{equation}
% q(\chi) = \int_0^\infty d\lambda e^{\vartheta \lambda - \lambda^2/2}/\lambda.
% \end{equation}
